### 2018-02-01 submission/manuscript.Rmd
discussion edits

Our observations build upon previous studies [@taq_Gohl_2016] by showing that even different HiFi DNA polymerases have significant differences in the number of OTUs and that changes in total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results, as well as the findings. Our observations show that HiFi DNA polymerase can have a noticeable affect on the OTUs generated and these differences are consistent across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index, based on cycle number, the data suggests that there may be differences between 15 and 20 cycles versus higher cycle numbers, such as 30x, that are commonly used. Additionally, there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). Conversely, we may not have enough samples to adequately judge the true magnitude difference between 20x and 25x communities and higher cycle numbers analyzed, even though there is a clear trend to suggest that 20x is very different then 25x [Figure 3]. This finding, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cycles are reached. Increasing the sub-sampling depth, for some DNA polymerase, may reduce some of these observed community differences at lower cycle numbers.

@@

Our observations suggest that there are clear HiFi DNA polymerase dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), the differences between the two mixes is not immediately apparent. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate.

>>>>

Our observations build upon previous studies [@taq_Gohl_2016] by showing that different HiFi DNA polymerases have both significant differences in the number of OTUs and that changes in total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results and conclusions. Our observations show that HiFi DNA polymerase can have a noticeable affect on the OTUs generated and these differences are consistent across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index the data suggests that there may be differences between 15 and 20 cycles versus higher cycle numbers, such as 30x, that are commonly used. There was few differences within individuals between corresponding 5-cycle increments (e.g. 15 to 20, 20 to 25, etc.). However, there is a clear trend that suggests that 20 cycles is very different then 25 cycles [Figure 3]. We may just not have enough samples to adequately assess the true magnitude difference between 20 and 25 cycle communities and higher cycle numbers analyzed. These findings, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cycles are reached. For some DNA polymerase, increasing the sub-sampling depth may reduce some of these observed community differences at lower cycle numbers.

@@

Our observations suggest that there are clear HiFi DNA polymerase dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A family polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), the differences between the two mixes is not immediately apparent. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate.

### 2018-02-01 submission/manuscript.tex
discussion edits

Our observations build upon previous studies (Gohl et al., 2016) by
showing that even different HiFi DNA polymerases have significant
differences in the number of OTUs and that changes in total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results, as well as
the findings. Our observations show that HiFi DNA polymerase can have a
noticeable affect on the OTUs generated and these differences are
consistent across sub-sampling depth and PCR cycle number {[}Figure
2-4{]}. These differences were observed in high biomass samples, where
biases introduced by such components like kit contamination may have
less of an effect, suggesting that these differences may be exacerbated
in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index,
based on cycle number, the data suggests that there may be differences
between 15 and 20 cycles versus higher cycle numbers, such as 30x, that
are commonly used. Additionally, there was no difference within
individuals between corresponding 5 cycle increments (e.g.~15 to 20, 20
to 25, etc.). Conversely, we may not have enough samples to adequately
judge the true magnitude difference between 20x and 25x communities and
higher cycle numbers analyzed, even though there is a clear trend to
suggest that 20x is very different then 25x {[}Figure 3{]}. This
finding, in conjunction with the PERMANOVA results, suggest that cycle
number can change bacterial community calculations but that these
differences are minimal once 25 cycles are reached. Increasing the
sub-sampling depth, for some DNA polymerase, may reduce some of these
observed community differences at lower cycle numbers.

@@

Our observations suggest that there are clear HiFi DNA polymerase
dependent differences in both per base error rate and chimeras that
cannot be removed using bioinformatic approaches {[}Figure 4 \& 5{]}.
Although it may be a natural assumption that the variation may be due to
the DNA polymerase family, the highest chimera rate, from Platinum, was
a family A polymerase while the lowest, from Accuprime, was also an A
polymerase (Ishino \& Ishino, 2014). In fact, from the material safety
data sheet (MSDS), the differences between the two mixes is not
immediately apparent. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate.

>>>>

Our observations build upon previous studies (Gohl et al., 2016) by
showing that different HiFi DNA polymerases have both significant
differences in the number of OTUs and that changes in total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results and
conclusions. Our observations show that HiFi DNA polymerase can have a
noticeable affect on the OTUs generated and these differences are
consistent across sub-sampling depth and PCR cycle number {[}Figure
2-4{]}. These differences were observed in high biomass samples, where
biases introduced by such components like kit contamination may have
less of an effect, suggesting that these differences may be exacerbated
in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index
the data suggests that there may be differences between 15 and 20 cycles
versus higher cycle numbers, such as 30x, that are commonly used. There
was few differences within individuals between corresponding 5-cycle
increments (e.g.~15 to 20, 20 to 25, etc.). However, there is a clear
trend that suggests that 20 cycles is very different then 25 cycles
{[}Figure 3{]}. We may just not have enough samples to adequately assess
the true magnitude difference between 20 and 25 cycle communities and
higher cycle numbers analyzed. These findings, in conjunction with the
PERMANOVA results, suggest that cycle number can change bacterial
community calculations but that these differences are minimal once 25
cycles are reached. For some DNA polymerase, increasing the sub-sampling
depth may reduce some of these observed community differences at lower
cycle numbers.

@@

Our observations suggest that there are clear HiFi DNA polymerase
dependent differences in both per base error rate and chimeras that
cannot be removed using bioinformatic approaches {[}Figure 4 \& 5{]}.
Although it may be a natural assumption that the variation may be due to
the DNA polymerase family, the highest chimera rate, from Platinum, was
a family A polymerase while the lowest, from Accuprime, was also an A
family polymerase (Ishino \& Ishino, 2014). In fact, from the material
safety data sheet (MSDS), the differences between the two mixes is not
immediately apparent. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate.
